---
title: "Other quantitative methods"
author: "Macartan Humphreys"
numbersections: true
header-includes:
  - \usepackage{amsmath, amssymb, bbm, amstext, array, listings, mathtools, multicolumn, caption, color, graphics, ulem, caption, changepage, atbegshi, soul, xcolor, CasualQueries}
  - \newcommand\E{\mathbb{E}}
  - \newcommand\V{\mathbb{V}}
  - \hypersetup{colorlinks=true,linkcolor=red}
  - \usepackage{ulem}
  - \pdfstringdefDisableCommands{\let\sout\relax}
fontsize: 11pt  
output:
  beamer_presentation:
    slide_level: 2
    keep_tex: false
    includes:
      in_header: include_nav.txt
  ioslides_presentation:
    self_contained: yes
    widescreen: no
---

## Recap: Differences in means worked when we had randomization

* We want to estimate: $$\frac1n\sum_i((Y_i(1)) - Y_i(0))$$

* We estimate using: $$\frac1{n_t}\sum_{i \in T}Y_i - \frac1{n_c}\sum_{i \in C}Y_i$$

* This works because, with randomization $\frac1{n_t}\sum_{i \in T}Y_i =  \frac1n\sum_i((Y_i(1))$ *in expectation* -- that is, on average the sample average is the population average. Similarly $\frac1{n_c}\sum_{i \in C}Y_i =  \frac1n\sum_i((Y_i(0))$ *in expectation*.


## But in the absence of randomization there are inferential risks

Difficulties once assignment is related to potential outcomes.



```{r, echo = FALSE, out.height="50%", fig.width=2, fig.height=1, fig.align='center', fig.cap = "Here X might be related to Y even though it does not cause Y"}
plot(make_model("X<-W -> Y"))
```


## In fact in the absence of randomization a model is *required* 

* Randomization is **not required** for causal inference.

* But without it you need some alternative argument for why your estimates from the treatment group capture *what would occur in the control group* if they were treated (and vice versa)

* What's more you will need a **model**:

    * Some variables might *have* to be taken into account in order to ensure no confounding
    * Some variables might *have* to *not* be taken into account in order to ensure no confounding
    * Sometimes these might be the same variables!   
    

## In the absence of randomization a model is *required* 

This is shown by Pearl (1995):

```{r, echo = FALSE, fig.width = 6, fig.height = 4,  fig.align="center", out.width='.9\\textwidth'}
source("../code/hj_dag.R")
par(mar=c(1,1,1,1))
hj_dag(x = c(-1, 0, -1, 1, 3),
       y = c(0, 2, 4, 1, 2),
       names = c( expression(paste(U[2])),
         "Z", expression(paste(U[1])), "X", "Y" ),
       arcs = cbind( c(1, 3, 1, 3, 4, 2),
                     c(2, 2, 4, 5, 5, 4)),
       title = "Z is a confound but controlling for it can induce bias",
       padding = .4, contraction = .15) 

```


## Alternatives

Randomization provides a very useful benchmark. Other strategies seek to approximate the magic of randomization:

* **Experimental Control** seeks to ensure that *ceteris* is *paribus* (induced unit homogeneity)
* **Adjustments**. Controlling: Regression,  Matching and Weighting
* **Instrumental variables** or **Natural experiments**---seeks a shock that approximates randomization. 
* **Difference in differences**---assumes that once you account for common time trends cases are as-if randomized
* **Regression discontinuity**---assumes that cases are as-if randomized *around a threshold* (there are also motivations for RDD that do not assume as-if randomization) 
* **Synthetic Matching** and other model based approaches

## Adjustment methods Intuition

Key idea is to figure out effects *conditional* on the values others nodes my take.

Our problem:


```{r, echo = FALSE, out.height="50%", fig.width=2, fig.height=1, fig.align='center'}
library(CausalQueries)
plot(make_model("X<-W -> Y <- X"))
```



Our solution:

$$X \rightarrow Y \>\>\> \> (\text{given }W = 0)$$
$$X \rightarrow Y \>\>\> \> (\text{given }W = 1)$$
We can estimate effects *within* similar sets and then average the results (weighting by the size of the sets) 

## RDD Intuition:

e.g. compare units in which the margin of victory was 1 vote for democrats against those for which it was -1. We expect that these are approximately identical, on average, is *all* regards and get an estimate of the effect of victory on some outcome *at the threshold*

```{r, echo  = FALSE, out.width="66%"}

x = seq(-1, 1, .01)
y = x+ rnorm(length(x))/10 + (x<0)/2 + .4*(x<0)*x
plot(x,y, xlab = "x: running variable (treatment if x>0)")
abline(a = 0, b = 1.1, col = "lightgreen")
abline(a = 0.5, b = 1.4, col = "lightblue")
```




## Diff-in-diff Intuition:

* examine the difference between treated and control in the before-after difference
* OK under the assumption that the change in the control group is the same as what the change would have been in the treatment group absent treatment

```{r, echo  = FALSE, out.width="66%"}

df <- fabricate(N =2,
          date = c(-1, 1),
          y_control = date + 1,
          y_treatment = date + 2 + (date==1)*1,
          y_control_counterfactual = date + 2) %>%
  gather("var", "value", - date, -ID)


ggplot(df, aes(date, value, color = var)) + geom_line()  + geom_point()

```

