---
title: "Causal inference and experimental methods"
author: "Macartan Humphreys"
date: Feb 2017
numbersections: true
header-includes:
  - \usepackage{amsmath, amssymb, bbm, amstext, array, listings, mathtools, multicolumn, caption, color, graphics, ulem, caption, changepage, atbegshi, soul}
  - \newcommand\E{\mathbb{E}}
  - \newcommand\V{\mathbb{V}}
  - \hypersetup{colorlinks=true,linkcolor=red}
  - \usepackage{ulem}
  - \pdfstringdefDisableCommands{\let\sout\relax}
fontsize: 11pt  
output:
  beamer_presentation:
    slide_level: 2
    keep_tex: false
    includes:
      in_header: include_nav.txt
  ioslides_presentation:
    self_contained: yes
    widescreen: no
---
  
```{r, echo=FALSE, warnings = FALSE, message = FALSE, include = FALSE}
rm(list = ls())
set.seed(343)
suppressMessages({
  library(DeclareDesign)
  library(xtable)
  library(knitr)
  library(sandwich)
  library(lmtest)
  library(ggplot2)
  library(dplyr)
  library(knitr)
  library(randomizr)

})
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = TRUE, message = FALSE, 
                      eval = TRUE, tidy = FALSE, tidy.opts = list(width.cutoff = 50),
                      strip_white = FALSE, results = 'asis', fix.ext = 'pdf', cache = TRUE)

```


```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Sampling and statistical error

## Sampling variability

Say that we had a world that had potential outcomes like this:

```{r, echo = FALSE, comment = " "}
set.seed(1)
n <- 10
Y0 <- sample(c(0,0,0,0,0,0,1,1,1,1))
Y1 <- sample(c(0,0,0,1,1,1,1,1,1,1))
```

```{r, echo = FALSE}
kable(cbind(ID = 1:n,Y0,Y1))
```

**Question**: What is the average treatment effect in this world?
**Answer**: 


## Sampling variability

Say that we had a world that had potential outcomes like this:

```{r, echo = FALSE, results ='asis'}
kable(cbind(ID = 1:n, Y0,Y1))
```

**Question**: What is the average treatment effect in this world?
**Answer**: Average(Y1) - Average(Y0) = `r mean(Y1 - Y0)`

## What might we find from an experiment?

Say we select a sample of units and look at their treatment outcomes, and another sample to look at their control outcomes.
```{r, comment = " ", echo = TRUE}
Z <- complete_ra(n)  # A random assignment
Y <- Y1*Z + Y0*(1-Z) # The outcome given this assignment
```

```{r, echo = FALSE}
kable(data.frame(ID = 1:n, Y0,Y1, Z, Y))
```

## What might we find from an experiment?

Estimate, caculated as difference-in-means:

```{r, echo = TRUE} 

mean(Y[Z==1]) - mean(Y[Z==0])

```

Or, using regression:

```{r, echo = TRUE, eval =FALSE} 
lm(Y~Z)
```

```{r, echo = FALSE} 
kable(coef(summary(lm(Y~Z))), digits = 2)
```

## What might we find from an experiment?

Now imagine doing this many times, with a different $Z$ each time:

```{r, echo = FALSE}

b <- NA

Z <- complete_ra(n)

```

```{r}
kable(rbind(Z, Y = Y1*Z + Y0*(1-Z)), col.names = 1:n)
```

With a different allocation each time we get a difference estimated effect each time: `r mean(Y1[Z==1] - Y0[Z==0])`

## What might we find from an experiment?

Plotted:

```{r, warning = FALSE, out.width="60%", fig.align='center', echo = FALSE}
f <- function(b, main = NULL, ylim = c(0.5, 3)) stripchart(b, method="stack", offset=0.5, pch=1, xlim = c(-1, 1), ylim = ylim, main = main)
b <- c(b, mean(Y1[Z==1] - Y0[Z==0]))

maintitle = "Possible estimates"
maxs <- 20
s   <- 0
```

```{r, echo = FALSE}
f(b, main = maintitle)
```


```{r child="sampling_variability.Rmd", echo=FALSE}
```

```{r, echo = FALSE}
b1 <- b
b <- NA
Y1 <- Y0
s   <- 0
maintitle <- "Possible estimates | Null"
```

# What might we find if there were no effect?


```{r child="sampling_variability.Rmd", echo=FALSE}
```


## Compare

```{r, echo = FALSE}
par(mfrow = c(2,1))
f(b, main = "Null")
f(b1,main =  "Expected")
```

## How would this look with more data?

```{r, echo = FALSE}
n <- 50
Y0 <- rbinom(n, 1, p = .4)
Y1 <- rbinom(n, 1, p = .7)
Zs <- replicate(1000, complete_ra(n))
b0 <- sapply(1:ncol(Zs), function(i) mean(Y0*(Zs[,i]==1) - mean(Y0*(Zs[,i]==0))))
b1 <- sapply(1:ncol(Zs), function(i) mean(Y1*(Zs[,i]==1) - mean(Y0*(Zs[,i]==0))))

par(mfrow = c(2,1))
hist(b0, xlim = c(-1,1)/2, main = "Null: 50"); box()
hist(b1, xlim = c(-1,1)/2, main = "Expected: 50"); box()

```

## How would this look with more data?

```{r, echo = FALSE}
n <- 200
Y0 <- rbinom(n, 1, p = .4)
Y1 <- rbinom(n, 1, p = .7)
Zs <- replicate(1000, complete_ra(n))
b0 <- sapply(1:ncol(Zs), function(i) mean(Y0*(Zs[,i]==1) - mean(Y0*(Zs[,i]==0))))
b1 <- sapply(1:ncol(Zs), function(i) mean(Y1*(Zs[,i]==1) - mean(Y0*(Zs[,i]==0))))

par(mfrow = c(2,1))
hist(b0, xlim = c(-1,1)/2, main = "Null: 200"); box()
hist(b1, xlim = c(-1,1)/2, main = "Expected: 200"); box()
```

## Power

You can read statistical power right off these graphs. 

* From the top graph you can see the probability of getting any type of effect, under the null. 
* From the bottom graph you can see the probability of getting any such effect

```{r, warning = FALSE, out.width="70%", fig.align='center', echo = FALSE}

par(mfrow = c(2,1))
hist(b0, xlim = c(-1,1)/2, main = "Null: 200"); box();  abline(v = quantile(b0, .975), col = "red")

hist(b1, xlim = c(-1,1)/2, main = "Expected: 200"); box(); abline(v = quantile(b0, .95), col = "red")

text(.35, 40, paste("Power = ", mean(b1 >= quantile(b0, .975))))
```

## Power

More generally this power depends on:

* How many units you have
* How big the effect is
* How noisy the process is

For a clustered design, power depends on how many clusters there are and how similar units are within clusters

* In general the more clusters the better
* The more dissimilar units are within clusters the better

Power also depends on variability from randomization. Blocking helps.

* In general the more blocks the better the better
* The more similar units are within blocks the better


## Small Effects

```{r, warning = FALSE, out.width="80%", fig.align='center'}
library(pwr)
p.t.two <- pwr.t.test(d=.1, power=0.8, type="two.sample", alternative="two.sided")
plot(p.t.two)
```

## Bigger effects

```{r, warning = FALSE, out.width="80%", fig.align='center'}
library(pwr)
p.t.two <- pwr.t.test(d=.3, power=0.8, type="two.sample", alternative="two.sided")
plot(p.t.two)
```


## Power Calculators

There are many power calculators.  

* Most are quite rudimentary. 
* For cluster designs with binary or continuous outcomes, see \url{https://egap.shinyapps.io/Power_Calculator/}
* But you can simulate *any* design and estimate power for any quantity

## Power Calculators \url{https://egap.shinyapps.io/Power_Calculator/}


\includegraphics[width=17.2cm]{calc.png}
