---
title: "Your title here"
subtitle: subtitle here
format:
  pdf:
    number-sections: true
    toc: false  # to done manually
    papersize: a4
    keep-tex: false
date: "`r format(Sys.Date(), '%d %B %Y')`"
author: "Your Name Here"
thanks: \texttt{email@address}
abstract: This is the abstract text.  Give the question, why it is important, your strategy, and your headline conclusions. All up front. For a thesis the abstract can be slightly longer. Abstracts can, but typically do not, include citations.
bibliography: assets/bib.bib
---

```{r setup, include = FALSE}
pacman::p_load(DeclareDesign,  knitr, plotly, tidyverse, texreg)
load("saved/analysis.Rdata")
theme_set(theme_bw())

# a helper to avoid double captions when using texreg
tex_clean <- function(models, ...) {
  out <- texreg::texreg(models, ..., return.string = TRUE)
  # Remove \caption{} and \label{} lines
  out <- gsub("\\\\caption\\{.*?\\}", "", out)
  out <- gsub("\\\\label\\{.*?\\}", "", out)
  return(out)
}

```



\newpage


# Introduction {#sec-intro}

* The question
* Why it is important
* How you answer it
* What you find
* How the thesis is structured


Some people use commenting in the text to ensure that each paragraph has a purpose

<!-- question -->

Why is it that...

<!-- strategy -->

To address this question I ...


Note: Introductions conventionally do not have subsections, or figures, or tables. 

# Theory

Describe literature that sheds light on why your project is important, clarifying what is known or not

## Cross Referencing

* To reference a figure with example label "plot", use  e.g.  `@fig-plot`.  @fig-1
* Analogously, to reference a table with example label "data", use  e.g.  `@tbl-data`.  @tbl-simple.
* To reference a section, such as the Introduction (@sec-intro), use  e.g.  `@sec-intro`.
* To reference an equation, same (@eq-euclid), use e.g. `@eq-euclid`.

For complete information on cross referencing with Quarto, see <https://quarto.org/docs/authoring/cross-references.html>.

## Citations

`.bib` information is sitting in `assets/bib.bib` 

* For a citation in parentheses use `[@putnam2000bowling]` and for a text citation: `@putnam2000bowling`. 
* These render as [@putnam2000bowling] and @putnam2000bowling

Multiple citations can be given as `[@putnam2000bowling;@blair2023research]`, producing  [@putnam2000bowling;@blair2023research]


# Design

## Case

Why you are working on the cases you are working on

## Data sources

where did you get the data?

## Data patterns

show the data!

## Analysis strategy

What strategies you will use

You can use \LaTeX natively. See  @eq-euclid.

$$
f(x) = -\frac{1}2(1-x)^2
$$ {#eq-euclid}



# Main results

# A deeper dive

Chapter could be:

1. second analysis
2. case studies
3. alternative interpretations
4. mechanisms
5. heterogeneity
6. robustness


## A sample figure

Here is figure. See @fig-1.

```{r, fig-1}
#| echo: false
#| fig-cap: "A sample figure: Information about this figure"

figure_1

```


## A table

See @tbl-simple. 


```{r}
#| label: tbl-simple
#| tbl-cap: "This is my table"
#| echo: false
#| results: 'asis'
#| 

tex_clean(models, include.ci = FALSE)

```



# Discussion


# Conclusion

* Recap of what was learned
* Caveats
* Implications: for policy, for research
* (Perhaps: Next steps)


Note: Conclusions conventionally do not have subsections, or figures, or tables. 

\newpage

# References

<!-- the pattern below controls the placement of the references -->

:::{#refs}
:::

\newpage

\appendix

# Appendixes {-}

# Appendix section

Details of the first appendix section.

## Appendix subection 

text

## Another subsection

text

# More

Text