---
title: Article Template
format:
  nature-pdf:
    # template credit: https://github.com/christopherkenny/nature
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
#   classoption:
#      - lineno
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: citeproc
    keep-tex: true
    number-sections: true
    embed-resources: true
    include-in-header:  
        - text: |
            \usepackage{multicol}
            \usepackage{xcolor}
            \usepackage{appendix}
            \usepackage{lipsum}
  html:
    code-fold: true
    html-math-method: katex
  docx: default
# The `csl` style will be used with LaTeX-based rendering of`nature-pdf`
# if cite-method is `citeproc`. This is also the default if cite-method
# is not specified.
csl: _extensions/nature/csl/springer-basic-author-date.csl
author:
  - name: Author One
    affiliations:
      - name: Humboldt University
        id: 1
        department: Social Sciences
        address: street
        city: Berlin
        state: Berlin
        postal-code: 12345
      - name: WZB
        id: 2
        department: IPI
        address: here
        city: Berlin
        state: Berlin
        postal-code: 12345
    attributes:
      equal-contributor: true
  - name: Author Two
    affiliations:
      - name: Humboldt University
        id: 1
        department: Social Sciences
        address: street
        city: Berlin
        state: Berlin
        postal-code: 12345
    email: corresponding@email.com
    attributes:
      equal-contributor: true
      corresponding: true
abstract: |
  The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted. 
keywords: [template, demo]
bibliography: bib.bib
appendix: true

---

```{r, include = FALSE}
# Commands that are useful for extracting code and replicating results:
# knitr::purl("my_paper.qmd")
# knitr::spin("my_paper.R")

# A tool to help with citations
# devtools::install_github("crsh/citr")

# Packages
if (!require(pacman)) install.packages("pacman")
pacman::p_load(
  tidyverse,
  DeclareDesign,
  knitr,
  kableExtra,
  modelsummary)
```
\newpage

# Introduction {#sec-intro}

This template is based on the quarto template provided by Christopher Kenny.
Note a range of different templates are available with various options: consult  https://github.com/christopherkenny/nature

1. Your question. Why we care
2. Your strategy, findings and the contribution
3. Roadmap

Some people use commenting in the text to ensure that each paragraph has a purpose

`# question`

Why is it that...

`# strategy`

To address this question I ...

`# etc`

\lipsum[1-3] 

# Motivation: Theory, Literature  {#sec-motivation}

Describe literature that sheds light on why your project is important, clarifying what is known or not

You can use \LaTeX natively. See  @eq-euclid.

$$
f(x) = -\frac{1}2(1-x)^2
$$ {#eq-euclid}

\newpage

# Results

Tables and Figures can be produced on the fly or imported.

## A sample figure

Here is figure produced on the fly. See @fig-1.
Code is shown but you would hide this by setting `echo: false`

```{r, fig-1}
#| echo: true
#| fig-cap: "sample figure"

read_rds("data/data.rds") |>
  ggplot(aes(X, Y)) + geom_point()

```

\newpage


## A sample summary stats table

Produced on the fly. See @tbl-simple.


```{r}
#| label: tbl-simple
#| echo: true

read_rds("data/data.rds") |>
  summarize(`mean Y` = mean(Y), 
            `mean X` = mean(X), 
            correlation = cor(X, Y)) |>
  kable(digits = 2, caption = "A sample table")

```
\newpage

## Analysis in parallel

Produced on the fly. 

```{r analysis}
#| label: analysis
#| echo: true
#| 
# generate a list of models
two_models <- 
  list(
  lm_robust(Y~1, data = read_rds("data/data.rds")),
  lm_robust(Y~X, data = read_rds("data/data.rds")))


```

\newpage

## Analysis output


Then print. Here using `modelsummary`. Lots of scope for customization: https://modelsummary.com/articles/appearance.html


```{r tbl-ols}
#| label: tbl-ols
#| echo: true
#| results: asis

# print nicely
two_models|>
  modelsummary(caption = ": some analyses", note = "some notes") 

```
\newpage

# Discussion: Referencing and cross referencing

## Cross Referencing

* To reference a figure with example label "plot", use  e.g.  `@fig-plot`.  @fig-1
* Analogously, to reference a table with example label "data", use  e.g.  `@tbl-data`.  @tbl-simple.
* To reference a section, such as the Introduction (@sec-intro), use  e.g.  `@sec-intro`.
* To reference an equation, same (@eq-euclid), use e.g. `@eq-euclid`.

For complete information on cross referencing with Quarto, see <https://quarto.org/docs/authoring/cross-references.html>.

## Citations

* For a citation in parentheses use `[@putnam2000bowling]` and for a text citation: `@putnam2000bowling`. 
* These render as [@putnam2000bowling] and @putnam2000bowling

Multiple citations can be given as `[@putnam2000bowling;@blair2023research]`, producing  [@putnam2000bowling;@blair2023research]

\newpage

# Conclusion

* Recap of what was learned
* Caveats
* Implications: for policy, for research
* (Perhaps: Next steps)


\newpage

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::

\newpage

\appendix

# Appendixes {-}

# Appendix section

Details of the first appendix section.

## Appendix subection 

text

## Another subsection

text

# More

Text