---
title: Article Template
format:
  nature-pdf:
    # template credit: https://github.com/christopherkenny/nature
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
#   classoption:
#      - lineno
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: citeproc
    keep-tex: true
    number-sections: true
    embed-resources: true
    include-in-header:  
        - text: |
            \usepackage{multicol}
  html:
    code-fold: true
    html-math-method: katex
  docx: default
# The `csl` style will be used with LaTeX-based rendering of`nature-pdf`
# if cite-method is `citeproc`. This is also the default if cite-method
# is not specified.
csl: _extensions/nature/csl/springer-basic-author-date.csl
author:
  - name: Author One
    affiliations:
      - name: Humboldt University
        id: 1
        department: Social Sciences
        address: street
        city: Berlin
        state: Berlin
        postal-code: 12345
      - name: WZB
        id: 2
        department: IPI
        address: here
        city: Berlin
        state: Berlin
        postal-code: 12345
    attributes:
      equal-contributor: true
  - name: Author Two
    affiliations:
      - name: Humboldt University
        id: 1
        department: Social Sciences
        address: street
        city: Berlin
        state: Berlin
        postal-code: 12345
    email: corresponding@email.com
    attributes:
      equal-contributor: true
      corresponding: true
abstract: |
  The abstract serves both as a general introduction to the topic and as a brief, non-technical summary of the main results and their implications. Authors are advised to check the author instructions for the journal they are submitting to for word limits and if structural elements like subheadings, citations, or equations are permitted.
keywords: [template, demo]
bibliography: bibliography.bib

---

```{r, include = FALSE}
library(tidyverse)
library(DeclareDesign)
library(knitr)
library(modelsummary)
```
\newpage

# Introduction {#sec-intro}

Your question. why we care

Your findings and the contribution

Roadmap


# Motivation: Theory, Literature  {#sec-motivation}


## Cross Referencing

* To reference a figure with example label "plot", use `@fig-plot`.  @fig-1
* Analogously, to reference a table with example label "data", use `@tbl-data`.  @tbl-simple.
* To reference a section, such as the Introduction (@sec-intro), use `@sec-intro`.

For complete information on cross referencing with Quarto, see <https://quarto.org/docs/authoring/cross-references.html>.

## Citations

* For a citation in parentheses use `[@greenwade93]` and for a text citation: `@greenwade93`. 
* These render as [@greenwade93] and @greenwade93

Multiple citations can be given as `[@greenwade93;@knuth1984texbook]`, producing  [@greenwade93;@knuth1984texbook]

\newpage

# Tables and Figures

These can be produced on the fly or imported.

## A sample figure

Produced on the fly. See @fig-1.

```{r, fig-1}
#| echo: true
#| fig-caption: "sample figure"

read_rds("data/data.rds") |>
  ggplot(aes(X, Y)) + geom_point()

```

\newpage


# A sample summary stats table

Produced on the fly. See @tbl-simple.


```{r}
#| label: tbl-simple
#| echo: true

read_rds("data/data.rds") |>
  summarize(`mean Y` = mean(Y), 
            `mean X` = mean(X), 
            correlation = cor(X, Y)) |>
  kable(digits = 2, caption = "A sample table")

```
\newpage

# A sample analysis table

Produced on the fly. 

```{r tbl-ols}
#| label: tbl-ols
#| echo: true
#| results: asis

list(
  lm_robust(Y~1, data = read_rds("data/data.rds")),
  lm_robust(Y~X, data = read_rds("data/data.rds")))|>
  modelsummary()

```

\newpage

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::
