---
title: "Writing: some principles"
author: "Macartan Humphreys"
date: "10/5/2020"
output: beamer_presentation
bibliography: bib.bib
urlcolor: orange
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Structure

Notes here: http://macartan.nyc/teaching/how-to-write/


# Front matter

* Include a  **table of contents**. 
* You can include **acknowledgments** and thank colleagues and friends---anyone who gave you support
* You can  **dedicate** this to someone if you like
* You can have both a short abstract (quarter page) and a longer executive summary (two pages). Include:

  * brief motivation
  * brief strategy 
  * main findings : positive or negative

# Table of contents

* This should be automated
* In Word: use Styles to select formats as "Header 1" "Header 2" etc; then References / Insert Table of Contents
* In  latex using `\section{}` and `subsection`
* In R markdown using #, ## 

Number sections e.g.  1, 1.1, 1.12, 1.2, 2.



# References

* The bibliography should contain an entry for every work cited---including websites---and should contain entries *only* for work cited.
* There are *lots* of rules around correct  formatting in the text and at the end. Plus there are different sets of rules.
* You should get this right but you shouldn't lose time.

References in sentences either have the year in parentheses or the year and last name in parentheses. They should include page references when possible.

Enter like this:

* `@putnam2000bowling` said some great stuff
* Putnam said some great stuff `[@putnam2000bowling]`
* `@putnam2000bowling [p. 7]` said some great stuff
* Putnam said some great stuff `[@putnam2000bowling, p. 7]`

[More](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)

# References

* The bibliography should contain an entry for every work cited---including websites---and should contain entries *only* for work cited.
* There are *lots* of rules around correct  formatting in the text and at the end. Plus there are different sets of rules.
* You should get this right but you shouldn't lose time.

References in sentences either have the year in parentheses or the year and last name in parentheses. They should include page references when possible.

Displays like this:

* @putnam2000bowling said some great stuff
* Putnam said some great stuff [@putnam2000bowling]
* @putnam2000bowling [p. 7] said some great stuff
* Putnam said some great stuff [@putnam2000bowling, p. 7]

[More](https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html)

# Grabbing references

I pull from google scholar mostly

* I do a search. e.g.  https://scholar.google.de/scholar?hl=en&as_sdt=0%2C5&q=putnam+bowling+alone&btnG=
* Then select the [bib](https://scholar.googleusercontent.com/scholar.bib?q=info:nQoGffkGw4MJ:scholar.google.com/&output=citation&scisdr=CgVA0Gv4EMHY0N6r9Ww:AAGBfm0AAAAAYA6u7WxVGiL8mR2PhobB31vo77xrrm3e&scisig=AAGBfm0AAAAAYA6u7eMD3UwBA3Q4An8W5YtfQrP8GTO6&scisf=4&ct=citation&cd=-1&hl=en) reference.
* And save that into a `.bib` text file


```{r, eval = FALSE, echo = TRUE}
@incollection{putnam2000bowling,
  title={Bowling alone},
  author={Putnam, Robert D},
  booktitle={Culture and politics},
  pages={223--234},
  year={2000},
  publisher={Springer}
}
```

# Web entries

@NYT

```{e, echo = TRUE, eval = FALSE}
@online{NYT,
  author = {{New York Times Editorial Board}},
  title = {Hong Kong Crackdown Is an Early Test for Biden},
  year = 2021,
  url = {https://www.nytimes.com/2021/01/24/opinion/hong-kong-arrests.html},
  urldate = {2021-01-25}
}
```

# Grabbing references

Other tools include Zotero, Endnote, Mendeley

There's a guide here: https://subjectguides.library.american.edu/c.php?g=479020&p=3323781

Principle is that you should spend a little time figuring out how to make this work and then not spend much time on it.

# Footnotes

* Footnotes generally preferred to endnotes
* Use footnotes sparingly
* Put singly at the end of a sentence, after the period.`^[Like this]`
* Put singly at the end of a sentence, after the period.^[Like this]
* Not like^[No!] that, or like this^[No].

# Writing

* Keep it sober, tight, straight to the point
* Keep formal: Avoid contractions (I've, would've)
* Avoid unnecessary superlatives
* Avoid opinion (absent evidence)



# Signposting

* Political science writing is not like literary writing
* Ordering: 
  * You don't build up and then reveal the findings at the end
  * You given then findings up front and then provide the evidence to support it
  * You should *not* assume that readers read linearly: they treat this as a compendium not a poem
  * So basically readers need to know what function every section has in the thesis and what function every paragraph has in a section

# Signposting: Lots please


**1. Introduction**
I show that natural resource abundance causes conflict. In section 2 I provide the logic. In section 3 I describe my strategy. Sections 4 and 5 give results and section 6 discusses implications.

**2. Theory**
Three theories predict an adverse effect of natural resource abundance. I describe each in turn, I then discuss strategies to distinguish between these accounts.

**3. Strategy**
I estimate the effect of natural resources on conflict using qualitative and quantitative strategies. I describe each in turn.
...

  
**5.1 Sierra Leone Case study**
I first present general background about this case, I then explore whether there is evidence in support of each of mechanisms 1, 2, and 3.



# Signposting: Theory

* Signposting is as important for theory theses
* The introduction should give the main argument and the arc of the evidence
* Chapters should, broadly, begin with a statement of their role and end with a conclusion



# Case study chapters

Be clear why you have a case study and what you want to do with it

a. Quick summary of why you have the case and what you learn from it
b. Justify case selection: why this case? How does it relate to other possible cases?
c. Say what you are looking for in the case and what you will infer depending on what you find
d. Describe sources
e. General description of the case
f. Specific findings relevant for theories
g. Case conclusions

# Quantitative results chapters

A standard ordering (whether in one chapter or many) is:

1. Describe hypotheses
2. Describe measures
3. Describe tests
4. Describe core results
5. Interpret results substantively
6. Describe robustness
7. Describe any extensions
8. Draw overall inferences

If divided into chapters you might have 1-3 in one chapter, 4-5 in a second, 6-7 in a third and 8 in a conclusion.


# Stay close to the data

* Display your raw data

```{r}
x = c(runif(10), 10)
y = c(runif(10) - x[1:10], 10)
plot(x,y)
abline(lm(y~x))

```

# The heart of an empirical analysis

* Commonly there is just table or figure at the *heart* of an empirical analysis
* Everything else is supporting the heart
* Know what the heart is and help readers find it quickly

* Often the  heart is a *figure* that pulls out the substantively important findings: the "quantities of interest"


# Tables and Figures

* When possible use figures rather than tables. 

* Tables and figures should usually be in the main body of the text, not at the end.

*  Tables and figures should be numbered and captioned/titled, in most cases. You can do automatic numbering word using “captions”, and in Latex using `\caption{}`

* Tables and other output should NOT be presented as Stata output or other raw output. There are many tools to produce nice output without a lot of work (`outreg2` in Stata, `stargazer` in `R`, but many more)



# Tables and Figures ctd

* Coefficients should be arranged in rows with comprehensible and consistent variable names

* standard errors should be shown in parentheses (or you should note what measure of uncertainty you are displaying if it is not standard errors)

* include descriptive statistics like number of observations (N) and R-squared should be included.

* Make sure to say what the dependent variables is

* Precision: numbers should normally reported to two significant digits. e.g.  0.12 not 0.000121313245

* Bottom line: It’s often useful and legitimate to give the bottom line of a table at the bottom of the table: “Table shows that there is no evidence that democracy causes growth”



# Mathematical writing

Use Latex when possible. 

* For instance: `$\frac{e^x}{1+x}$` renders as $\frac{e^x}{1+x}$
* This works the same way in Latex and `.Rmd`

* [This guide](https://www.google.com/url?sa=t&rct=j&q=&esrc=s&source=web&cd=&cad=rja&uact=8&ved=2ahUKEwjngpT9irfuAhVR-qQKHWmHBxcQFjAAegQIAhAC&url=http%3A%2F%2Fwww.math.ubc.ca%2F~cytryn%2Fteaching%2FscienceOneF10W11%2Fhandouts%2FOS.writingMathCommented.pdf&usg=AOvVaw18QDi0hKFxiPtZt160hEtE) to mathematical writing has some nice tips: 

# References

