---
output:
  word_document: default
  pdf_document:
    number_sections: yes
geometry: left=3.8cm,right=2.5cm,top=2.5cm,bottom=2.5cm
header-includes: \usepackage{amsmath} \usepackage{pdflscape}
bibliography: bib.bib
urlcolor: orange
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(stargazer)
```


\pagenumbering{roman}     

\newpage                  <!-- new page -->

\begin{center}        

\Huge{\sc A Great Thesis} % <!-- make large text -->

\bigskip                % <!-- blank lines -->
\bigskip
\bigskip

\large{Honors Student}

\bigskip
\bigskip
\bigskip
\bigskip
\bigskip

Thesis submitted to the Faculty of the Department of Political Science of Columbia University in the City of New York in partial fulfillment of the requirements for the
honors undergraduate program. 

\bigskip
\bigskip
\bigskip
1 April  2021

\end{center}              
\bigskip
\bigskip
\bigskip
\bigskip

**Advisory Committee:**

A
B
C


\newpage 

\begin{center}

\large{Abstract}

\end{center}

\bigskip

Multiple accounts suggest that online targeting advertising has substantial impacts on electoral outcomes. I study the effects of targeted campaigns by Oxford Analytica and find no evidence that their campaigns affected electoral outcomes in any district. Qualitative investigations suggest that the ads were either not seen or ignored by targetted voters.       

See formatting hints here: https://achale.gitlab.io/tutorialmarkdownthesis/


\newpage 

\tableofcontents


\newpage

\pagenumbering{arabic} 

# Introduction

Motivation

Relevant literatures

Overview of findings


I describe the design in section   \ref{design} and results in section \ref{results}

\newpage

# Design {#design}

## Inferential Strategy

Text

## Measurement strategy

Data, descriptive statistics

## Cases

Motivation, case selection

\newpage

# Results {#results}

## Main results 

See Figure \ref{fig:myfigure}.

```{r, fig.cap = "\\label{fig:myfigure} A figure with a detailed  note", echo = FALSE, out.width = "50%", echo = FALSE, fig.align = "center", fig.height = 6, , fig.width = 6}

plot(0:1, 0:1, type = "l")

```


See Table \ref{tab:mytab}

```{r mytab, echo = FALSE}

data.frame(name = c("A", "B"), value = 1:2) %>% kable(caption = "\\label{tab:mytab} A caption describing this table")

```


Regression results:

```{r, echo=FALSE, results='asis'}

df <- data.frame(Y = runif(10), X = runif(10), Z = runif(10)) 
M1 <- lm(Y~X, data = df)
M2 <- lm(Y~X +Z, data = df)

stargazer(M1, M2, header=FALSE)

```


## Intepretations

Text


## Robustness Extensions

\newpage

# Conclusion 

@putnam2000bowling said some great stuff

Putnam said some great stuff [@putnam2000bowling]

@putnam2000bowling [p. 7] said some great stuff

Putnam said some great stuff [@putnam2000bowling, p. 7]

See also @NYT.


\newpage

# References{-}
<div id="refs"></div>

\renewcommand{\thesection}{\Alph{section}}
\setcounter{section}{1}

\newpage
# Appendix{-}
## Model details {#app:model}
## Robustness results {#app:robust}

