---
title: "Merging"
date: "1/31/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

Here we have three data datasets and we merge the second and third into the first. Note that the first dataset has *two* sets of IDs (name and code). One of these is used to help merge with `df_2` and the other  with `df_3`.

# Basic merging

```{r}
df_1 <- data.frame(name = LETTERS[1:4], code = 1:4, X = runif(4))

df_2 <- data.frame(name = c("A", "B", "D"), var_2 = runif(3))

df_3 <- data.frame(code = 1:3, var_3= c("a", "v", "w"))

df <- df_1 %>% left_join(df_2) %>% left_join(df_3)

df %>% kable()
```


# WDI

Here is an example in which we extract a variable from WDI and then merge in 
region information.

```{r}

library(WDI)


# GDP data
df <- WDI::WDI(start = 1992, end = 1993)

head(df)

# Lets add in region and iso3c

# We make a codes dataset
country_codes <- WDI::WDI_data$country %>% data.frame() %>% 
  dplyr::select(iso2c,  iso3c, region)

head(country_codes)


# Merge in
df <- df %>% left_join(country_codes)

head(df)

# Lets get rid of the aggregates because Africa is not a country
df <- df %>% filter(region != "Aggregates")
head(df)

```



# Issues

Merging is in principle easy but in practice pretty hard. 

Key challenges:

## Different naming 

Merging will not work if you have different codes in teh datasets you are merging:

```{r}
df_1 <- data.frame(id = c("Germany", "U.K.", "Zaire"), X1 = 1:3)
df_2 <- data.frame(id = c("Germany", "UK", "DRC"), X2 = 5:7)

left_join(df_1, df_2)
```

We end up with missing data in X2 because the right ids were not found in the two datasets.

Better:

```{r}
df_1 %>% left_join(df_2)

df_1 %>% mutate(id = recode(id, "U.K." = "UK", "Zaire" = "DRC")) %>%
  left_join(df_2)
```

Best: have one dataset that has a complete list of authorative codes --- hopefully a very standard set of codes. Prep all other data so that they have the same coding system.  

## Incomplete data produces many holes


```{r}
df_1 <- data.frame(id = c("Germany"), X1 = 1)
df_2 <- data.frame(id = c("DRC"), X2 = 2)

left_join(df_1, df_2)
```

Better:


```{r}
df_0 <- data.frame(id = c("DRC", "Germany"))

df_0 %>% left_join(df_1) %>% left_join(df_2)
```

again: start off with a complete frame, with all countries, and add into this.


## Data is the wrong shape

Very often you want to combine two datasets but they have different shapes. Viz:

```{r}

df_1 <- data.frame(id = c("France", "Germany"), GDP_1990 = runif(2), GDP_1995 = runif(2))
df_1

df_2 <- data.frame(id = c("France", "France", "Germany", "Germany"), 
                   Year = c(1990, 1995, 1990, 1995),
                   inflation = runif(4))
df_2


```

we need to get these two into the  same shape in order to merge. I use `gather` here but there are many other reshape functions that might be better for your data.


```{r}
df_1 <- 
  df_1 %>% gather("Year", "GDP",  -id) %>%
  mutate(Year = gsub("GDP_", "", Year),
         Year = as.numeric(Year))

df_1 %>% left_join(df_2)
```

Notice that the joining was done on both id and Year since both of these were in the two datasets.
